Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-11-15 WSSSPE3/report/wg_main_industry_interaction.tex
Section 5.2 and Section 5.3: No content change
Section 5.3: tiny changes

Most scientific software is produced as a part of grant-funded research projects sponsored by federal governments. If we are interested in the sustainability of scientific software then we need to understand what exactly happens when that sponsorship ends. More than likely, the project and its resulting software will need to undergo some kind of transition in funding and consequently management.

There are a number of potential funding transitions that may occur:

\begin{itemize}
\item A project could be \textbf{refunded}, and development or maintenance of the software continues as planned.
\item A project might located a \textbf{new source of funding} in which case the software may be further developed or simply maintained as before. 
\item The project could transition to a \textbf{community supported model} whereby ownership, maintenance and stewardship of the software becomes similar to peer-production models in open-source (e.g., see Howison~\cite{howison_sustaining_2015}).
\item The project could receive some form of industry sponsorship in which case ownership of the intellectual property, licensing, maintenance activities, hosting, etc.\ may change significantly. 
\end{itemize}

@@

At WSSSPE3 our group was interested in better understanding successful pathways for scientific software to ``transition'' from government-funded research projects to industry sponsorship. 
(This may be an initially awkward phrase---some software projects will begin their life being sponsored by industry, or result in collaboration between industry and academia. 
In such cases, there is still a need to understand how IP and how maintenance of the software is sustained over time)

Although transitions are often studied under the broad umbrella of `technology transfer' we believe there are likely to be a number of different ways in which a pathway from initial production to long-term maintenance and secure funding is achieved. In short, industry sponsorship is an important aspect of sustaining scientific software, but our current understanding of these transitions focuses narrowly on commercial successes\slash failures.

@@

In looking at existing literature that addresses industry transitions many reports focus on benefits that accrue to the private sector, or to a government that originally sponsored the research project.
This literature does not address the impact that these transitions have on the accessibility, or usability of the software, or the impact that these transitions have on the career of the researchers involved.

Examples of the former scenario (benefits accruing to private sector) are as follows:
\begin{itemize}
\item REF Impact Case Studies: \url{http://impact.ref.ac.uk/CaseStudies/}
\item Background of projects funded in the UK: \url{http://gtr.rcuk.ac.uk/}
\item Dowling Review from the UK: addresses complexity of work between these two communities: \url{http://www.raeng.org.uk/policy/dowling-review}
\item Pathway to Impact - UK report: two pages of grant proposals are asked to forecast how they might impact (includes environmental, academic, economic)
\end{itemize}

@@

\textbf{GOAL}
\emph{What is the goal of doing research on transition pathways?}

@@

Complete a set of case studies which look at successful and unsuccessful transitions between academic researchers and industry.
This might address one of each of the transition types (described below).
Successful transitions are described as those that lead to either weak or strong sustainability (also defined below).

Create a generalizable framework that might allow for the study of different transition pathways (other than academia -> industry).

@@

We characterize transitions in the following ways:
\begin{itemize}
\item Handoff model: academia initially writes the software, industry (for profit or nonprofit) then takes over project. 
\item Co-Production Model: industry and academia interact throughout development of the project.
\item Sponsorship Model: academia writes and maintains the software, the industry contributes funding for the development\slash maintenance of software. In this example industry is also likely a user of the software.
\item Spinoff model: transition to for- or non-profit company owned by or in collaboration with original developers. 
\end{itemize}

We characterize sustainability in the following ways:
\begin{itemize}
\item Weak Sustainability: Software continues to be accessible, useful, and usable. 
\item Strong Sustainability: Software meets criteria above, but is also able to be reused for further innovation (i.e., issued non-restrictive open-source license).
\end{itemize}
See Becker et al.~\cite{Becker:2014} for extended discussion of weak vs.\ strong sustainability.
For more detail on the group's discussion, see Appendix~\ref{sec:appendix_industry_interaction}.

>>>>

Most scientific software is produced as a part of grant-funded research projects
sponsored by federal governments. If we are interested in the sustainability of
scientific software, then we need to understand what exactly happens when that
sponsorship ends. More than likely, the project and its resulting software will
need to undergo some kind of transition in funding and consequently management.

There are a number of potential funding transitions that may occur:  
%
\begin{itemize}

\item A project could be \textbf{refunded}, and development or maintenance of
the software continue as planned.

\item A project might locate a \textbf{new source of funding} in which case the
software may be further developed or simply maintained as before.

\item The project could transition to a \textbf{community supported model}
whereby ownership, maintenance, and stewardship of the software become similar
to peer-production models in open-source (e.g., see
Howison~\cite{howison_sustaining_2015}).

\item The project could receive some form of industry sponsorship in which case
ownership of the intellectual property, licensing, maintenance activities,
hosting, etc.\ may change significantly.

\end{itemize}

@@

At WSSSPE3 our group was interested in better understanding successful pathways
for scientific software to ``transition'' from government-funded research
projects to industry sponsorship. (This may be an initially awkward
phrase---some software projects will begin their life being sponsored by
industry, or result in collaboration between industry and academia. In such
cases, there is still a need to understand how IP and how maintenance of the
software is sustained over time.)

Although transitions are often studied under the broad umbrella of ``technology
transfer'', we believe there are likely to be a number of different ways in which
a pathway from initial production to long-term maintenance and secure funding are
achieved. In short, industry sponsorship is an important aspect of sustaining
scientific software, but our current understanding of these transitions focuses
narrowly on commercial successes\slash failures.

@@

In looking at existing literature that addresses industry transitions, many
reports focus on benefits that accrue to the private sector, or to a government
that originally sponsored the research project. This literature does not address
the impact that these transitions have on the accessibility or usability of the
software, or the impact that these transitions have on the career of the
researchers involved.

Examples of the former scenario (benefits accruing to private sector) are as
follows:
\begin{itemize}

\item REF Impact Case Studies: \url{http://impact.ref.ac.uk/CaseStudies/}

\item Background of projects funded in the UK: \url{http://gtr.rcuk.ac.uk/}

\item Dowling Review from the UK: addresses complexity of work between these two
communities: \url{http://www.raeng.org.uk/policy/dowling-review}

\item Pathway to Impact - UK report: two pages of grant proposals are asked to
forecast how they might impact (includes environmental, academic, economic) \choinote{say impact on what?}.

\end{itemize}

@@

\textbf{Goal}

\emph{What is the goal of doing research on transition pathways?}

@@

Complete a set of case studies which look at successful and unsuccessful
transitions of researchers between academia and industry. This might address 
each of the transition types (described below). Successful transitions are
described as those that lead to either weak or strong sustainability (also
defined below).

Create a generalizable framework that might allow for the study of different
transition pathways (other than academia to industry).

@@

We characterize transitions in the following ways:
\begin{itemize}

\item Handoff model: academia initially writes the software, industry (for-profit 
or nonprofit) then takes over the project.

\item Co-Production Model: industry and academia interact throughout development
of the project.

\item Sponsorship Model: academia writes and maintains the software; the
industry contributes funding for the development\slash maintenance of software.
In this example, industry is also likely a user of the software.

\item Spinoff model: transition to for- or non-profit company owned by or in
collaboration with original developers.

\end{itemize}

We characterize sustainability in the following ways:
\begin{itemize}

\item Weak Sustainability: Software continues to be accessible, useful, and
usable.

\item Strong Sustainability: Software meets criteria above, but is also able to
be reused for further innovation (i.e., issued non-restrictive open-source
license).

\end{itemize}
We refer readers to Becker et al.~\cite{Becker:2014} for extended discussion of weak versus strong
sustainability. For more detail on the group's discussion, see
Appendix~\ref{sec:appendix_industry_interaction}.
